### 2017-07-02 opsem.tex
Edits.

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). Certain terms that only appear at run-time are also
present in $c$. These include a \C{TXN} block tagged with sets ($\stl$
and $\stg$) of records that assume special meaning in operational
semantics, and a \C{FOREACH} loop that keeps track of the set ($s_1$)
of items already iterated, and the set ($s_2$) of items yet to be
iterated. Note that even the surface-level \C{FOREACH} shown here is a
little different from the one used in previous sections; its
higher-order argument has two arguments, $y$ and $z$, which are
invoked (during the reduction) with the set of already-iterated items,
and the current item, respectively. This form of \C{FOREACH} lends
itself to inductive reasoning, facilitating inductive proofs
(Sec.~\ref{sec:reasoning}).

@@

The top-level reduction relation

% Fundamental to our development is the notion of a trace invariant
% ($\I$). $\I$ is a function from traces ($\E$) to first-order logical
% formulas ($\Prop$) that define well-formedness
% constraints over traces.  The machine takes a step only if the
% resulting trace satisfies the constraints imposed by $\I$. This
% behaviour is captured by the auxiliary reduction rule
% \rulelabel{E-Aux} that factors out the trace extension aspect of the
% evaluation by abstracting away the operation-specific behaviour as a
% function that generates an appropriate effect. We let $\mathcal{F}$
% denote this function.  \rulelabel{E-Aux} uses $\mathcal{F}$ to
% generate a new effect and extend the trace ($\E = (\A,\visZ)$)
% \emph{only if} the well-formedness constraints imposed by $\I$ on $\E$
% (i.e., $\I(\E')$) are satisfied. Otherwise, it gets stuck. In an
% execution that runs to completion, every small-step preserves the
% well-formedness of a trace, thus ensuring the invariance of $\I$.
% Note that the semantics makes no assumptions about $\I$ other than its
% type. As such, it can be instantiated with any trace-parametric
% proposition that expresses constraints over the given trace. For
% instance, consider the $\psi_{RC}$ specification from
% \S\ref{sec:motivation}, but with bounded $T_1$ and $T_2$ instantiated
% with \C{Wd1} and \C{Wd2}, respectively. The instantiated specification
% is the following term:\vspace*{-8pt}

>>>>

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). The $\I$ in \C{TXN} block syntax is the
transaction's isolation specification, which will be explained later.
Certain terms that only appear at run-time are also present in $c$.
These include a \C{TXN} block tagged with sets ($\stl$ and $\stg$) of
records that assume special meaning in operational semantics, and a
\C{FOREACH} loop that keeps track of the set ($s_1$) of items already
iterated, and the set ($s_2$) of items yet to be iterated. Note that
even the surface-level \C{FOREACH} shown here is a little different
from the one used in previous sections; its higher-order argument has
two arguments, $y$ and $z$, which are invoked (during the reduction)
with the set of already-iterated items, and the current item,
respectively. This form of \C{FOREACH} lends itself to inductive
reasoning, facilitating inductive proofs (Sec.~\ref{sec:reasoning}).

@@

The top-level reduction relation defines small-step semantics of
transactions, and their parallel composition. A transaction always
comes tagged with an isolation specification $\I$ that dictates the
timing and nature of iterferences that the transaction can witness.
Formally, $\I$ is a function that maps a boolean ($\mathbb{B}$) and a
triplet of database states (three sets of records:
$\Pow(r)\times\Pow(r)\times\Pow(r)$) to a logical proposition
($\Prop$). The abstract machine calls $\I$ while reducing
$\ctxn{i}{\I}{c}$. The boolean argument identifies if the transaction
is still executing (\C{true}), or if it has finished execution and
ready to commit (\C{false}).  Among the triplet of sets (the second
argument), first is the (current) transaction-local database state
($\stl$), second is the state ($\stg$) of the global database when the
transaction last took a step, and third is the current state ($\stg'$)
of the global database.  Intuitively, $\stg'\neq\stg$ indicates an
interference from a concurrent transaction, and the $\I$ decides if
this interference is allowed or not, considering the local database
state ($\stl$). For instance, as described in
(Sec.~\ref{sec:motivation}), an RR transaction on Postgres defines
$\I$ as following:
\begin{smathpar}
\begin{array}{lcl}
\I\,\C{true}\,(\stl,\stg,\stg') & = & \stg' = \stg\\
\I\,\C{false}\,(\stl,\stg,\stg') & = & \forall(r\in\stl)(r'\in\stg).~
  r'.\idf = r.\idf \Rightarrow r'\in\stg'.
\end{array}
\end{smathpar}

The reduction of a $\ctxn{i}{\I}{c}$ begins by first converting it to
its run-time form $\ctxn{i}{\I,\stl,\stg}{c}$, where $\stl =
\emptyset$, and $\stg$ the current (global) database. The top-level
reduction rules in Fig.~\ref{fig:txnimp} assume this form. Rule
\rulelabel{E-Txn} reduces $\ctxn{i}{\I,\stl,\stg}{c}$ under a database
state ($\stg'$), only if the isolation specification ($\I$) allows the
interference between $\stg$ and $\stg'$ during the execution. Rule
\rulelabel{E-Commit} commits the transaction
$\ctxn{i}{\I,\stl,\stg}{c}$ by flushing its uncommitted records to the
database. This is done only if the interference between $\stg$ and
$\stg'$ is allowed during the commit by the isolation specification.

% For convenience, we define $\I_r = \I(\C{true})$ and $\I_c =
% \I(\C{false})$.
% Fundamental to our development is the notion of a trace invariant
% ($\I$). $\I$ is a function from traces ($\E$) to first-order logical
% formulas ($\Prop$) that define well-formedness
% constraints over traces.  The machine takes a step only if the
% resulting trace satisfies the constraints imposed by $\I$. This
% behaviour is captured by the auxiliary reduction rule
% \rulelabel{E-Aux} that factors out the trace extension aspect of the
% evaluation by abstracting away the operation-specific behaviour as a
% function that generates an appropriate effect. We let $\mathcal{F}$
% denote this function.  \rulelabel{E-Aux} uses $\mathcal{F}$ to
% generate a new effect and extend the trace ($\E = (\A,\visZ)$)
% \emph{only if} the well-formedness constraints imposed by $\I$ on $\E$
% (i.e., $\I(\E')$) are satisfied. Otherwise, it gets stuck. In an
% execution that runs to completion, every small-step preserves the
% well-formedness of a trace, thus ensuring the invariance of $\I$.
% Note that the semantics makes no assumptions about $\I$ other than its
% type. As such, it can be instantiated with any trace-parametric
% proposition that expresses constraints over the given trace. For
% instance, consider the $\psi_{RC}$ specification from
% \S\ref{sec:motivation}, but with bounded $T_1$ and $T_2$ instantiated
% with \C{Wd1} and \C{Wd2}, respectively. The instantiated specification
% is the following term:\vspace*{-8pt}
